Cooperative Kernels: GPU Multitasking for Blocking Algorithms

### 2017-07-06 write-ups/FSE/paper.tex
Done pass over Section 3.

Cooperative kernels are written using the following
extensions: $\transmit$, a qualifier on the variables of a
thread; $\offerkill$ and $\offerfork$, the key functions that enable
cooperative scheduling; and $\globalbarrier$ and $\mathsf{resizing\_global\_barrier}$
%$\resizingglobalbarrier$
primitives for inter-workgroup synchronisation.

@@

%\vspace{-1mm}
\myparagraph{Active Workgroups}
%
If the host application launches a cooperative kernel requesting $N$
workgroups, this indicates that the kernel should be executed with a
maximum of $N$ workgroups, and that as many workgroups as possible, up
to this limit, are desired.  However, the scheduler may initially
schedule fewer than $N$ workgroups, and as explained below the number
of workgroups that execute the cooperative kernel can change during
the lifetime of the kernel.

@@

%\vspace{-1mm}
\myparagraph{Global Barriers}
%
Because workgroups of a cooperative kernel are fairly scheduled, so a
global barrier primitive can be provided, and we specify two variants: $\globalbarrier$
and $\resizingglobalbarrier$.

@@

The semantics of $\resizingglobalbarrier$ can be modelled via calling $\offerfork$ and $\offerkill$
surrounded, and separated by calls to a $\globalbarrier$.
%
%\lstset{basicstyle=\small\tt,numbers=none}
%\begin{lstlisting}
%  $\globalbarrier()$;  $\offerfork$();
%  $\globalbarrier()$; $\offerkill()$;
%  $\globalbarrier()$;
%\end{lstlisting}
%\lstset{basicstyle=\scriptsize\tt,numbers=left}
%
The enclosing $\globalbarrier$ calls ensure that the change in number
of active workgroups from $M$ to $M'$ occurs entirely within the
resizing barrier, so that from a programmer's perspective, $M$ changes
atomically.  The middle $\globalbarrier$ ensures that forking occurs
before killing, so that workgroups $[0, \textrm{min}(M, M') - 1]$ are
left intact.

@@

The purpose of $\offerkill$ is to let the scheduler destroy a workgroup
in order to schedule higher-priority tasks.  The scheduler relies on the
cooperative kernel to execute $\offerkill$ sufficiently frequently that
soft real-time constraints of other workloads can be met.
%
Using our work stealing example: a workgroup offers itself to
the scheduler after processing each task.  If tasks are sufficiently
fast to process then the scheduler have ample opportunities to
de-schedule workgroups.  But if tasks are very time-consuming to
process then it might be necessary to rewrite the algorithm so that
tasks are shorter and more numerous, to achieve a higher rate of calls
to $\offerkill$.
%
Getting this non-functional requirement right is GPU- and
application-dependent.  In \mysec\ref{sec:sizingnoncoop} we conduct
experiments to understand the response rate that would be required to
co-schedule graphics rendering with a cooperative kernel, maintaining
a smooth frame rate.

>>>>

Cooperative kernels are written using the following
extensions: $\transmit$, a qualifier on the variables of a
thread; $\offerkill$ and $\offerfork$, the key functions that enable
cooperative scheduling; and $\globalbarrier$ and $\resizingglobalbarrier$
primitives for inter-workgroup synchronisation.

@@

%\vspace{-1mm}
\myparagraph{Active Workgroups}
%
If the host application launches a cooperative kernel requesting $N$
workgroups, this indicates that the kernel should be executed with a
\emph{maximum} of $N$ workgroups, and that as many workgroups as possible, up
to this limit, are desired.  However, the scheduler may initially
schedule fewer than $N$ workgroups, and as explained below the number
of workgroups that execute the cooperative kernel can change during
the lifetime of the kernel.

@@

%\vspace{-1mm}
\myparagraph{Global Barriers}
%
Because workgroups of a cooperative kernel are fairly scheduled, a
global barrier primitive can be provided.  We specify two variants: $\globalbarrier$
and $\resizingglobalbarrier$.

@@

The semantics of $\resizingglobalbarrier$ can be modelled via calling $\offerfork$ and $\offerkill$,
surrounded and separated by calls to a $\globalbarrier$.
%
%\lstset{basicstyle=\small\tt,numbers=none}
%\begin{lstlisting}
%  $\globalbarrier()$;  $\offerfork$();
%  $\globalbarrier()$; $\offerkill()$;
%  $\globalbarrier()$;
%\end{lstlisting}
%\lstset{basicstyle=\scriptsize\tt,numbers=left}
%
The enclosing $\globalbarrier$ calls ensure that the change in number
of active workgroups from $M$ to $M'$ occurs entirely within the
resizing barrier, so that $M$ changes atomically from a programmer's perspective.  The middle $\globalbarrier$ ensures that forking occurs
before killing, so that workgroups $[0, \textrm{min}(M, M') - 1]$ are
left intact.

@@

The purpose of $\offerkill$ is to let the scheduler destroy a workgroup
in order to schedule higher-priority tasks.  The scheduler relies on the
cooperative kernel to execute $\offerkill$ sufficiently frequently that
soft real-time constraints of other workloads can be met.
%
Using our work stealing example: a workgroup offers itself to
the scheduler after processing each task.  If tasks are sufficiently
fast to process then the scheduler will have ample opportunities to
de-schedule workgroups.  But if tasks are very time-consuming to
process then it might be necessary to rewrite the algorithm so that
tasks are shorter and more numerous, to achieve a higher rate of calls
to $\offerkill$.
%
Getting this non-functional requirement right is GPU- and
application-dependent.  In \mysec\ref{sec:sizingnoncoop} we conduct
experiments to understand the response rate that would be required to
co-schedule graphics rendering with a cooperative kernel, maintaining
a smooth frame rate.
